---
title: "Final Project EDA"
author: "Vaishali Devarakonda"
date: "Nov. 12, 2020"
output: html_notebook
---

#Intro

I will be examining the two data sources for this project: the first source is a dataset from Kaggle describing audience and critic ratings for a specific movie, created by user Mormukutchaudhary. This dataset by itself can have many uses. For example, moviegoers can examine this to determine whether a movie is worth watching. 

The other is a table from the-numbers.com that describes the market share for movie genres between 1995-2020. This dataset is more useful, for example, to people in the movie industry who are coming up with a new concept for a movie. It could also be useful for movie marketing teams to come up with a good strategy that could result in higher box office numbers and/or ratings

# Set Up

### Clean up environment and load packages
```{r}
# clean up the RStudio environment 
rm(list = ls())

# load all packages here: `mosaic`, `tidyverse`, and all others used
library(mosaic)
library(tidyverse)
library(tidyr)
library(rvest)
library(lubridate)
library(ggplot2)

#to import dataset I need
library(readr)

```

### Load the data you need
```{r}
Movie_Ratings <- read_csv("Movie Ratings.csv")
View(Movie_Ratings)
```

```{r}
#look at the first 6 rows
head(Movie_Ratings, 6)
```
Movie_ratings (from Kaggle) has the following case definition: "list the budget, audience, and critic ratings for a given movie".

```{r}
#rename columns for readability
Movie_Ratings <- Movie_Ratings %>% 
  rename(critic = 'Rotten Tomatoes Ratings %',
         audience = 'Audience Ratings %',
         year = 'Year of release')
```

The second data source comes from 'the-numbers.com'. However, I will be using a cleaned version of the data, which is loaded in a csv format.

```{r}
Genres <- read.csv("C:/Users/18vde/OneDrive/Desktop/PSU/STAT 184/Genres.csv")

Genres <- Genres %>% rename (rank = 'ï..Rank',
                             'BO_Rev (millions)'  = 'Total.Box.Office')

head(Genres, 6)
```
From this table, we can see that the case meaning is the following: "list the number of movies, box office revenue (in millions), number of tickets sold, and the market share (%) for each genre".


### Similarities between the data?

We can see that Movie_Ratings and Genres have a common variable: Genre. But how many values are common between the two data frames?
```{r}
Genres %>% distinct(Genre)
```

```{r}
Movie_Ratings %>% distinct(Genre)
```
As we can see, each genre from Movie_Ratings is present, which can lead us to ask the research question presented in the next section.

Let's look at these two data frames joined.
```{r}
MovieData <- Movie_Ratings %>% left_join(Genres, by = "Genre")

head(MovieData, 6)
```

# Research Question

After examining the data, a good research question that we could explore would be the following:

**Does the total Box Office Revenue generated by a genre have any influence on audience ratings for a particular film?**

While the data frames show little evidence of maintenance (Movie_Ratings was uploaded to Kaggle 2 years ago; Genres was only published recently), they are still very useful to answer this question. 

It may not always be the case that overall box office numbers will result in a comedy film being highly rated by viewers, for example. In addition, although ratings can be very subjective, there are often large groups of people who prefer certain genres and will rate its movies very highly.
