---
title: "Final Project EDA"
author: "Vaishali Devarakonda"
date: "Nov. 12, 2020"
output: html_notebook
---

#Intro

I will be examining the two data sources for this project: the first source is a dataset from Kaggle describing audience and critic ratings for a specific movie, created by user Mormukutchaudhary. This dataset by itself can have many uses. For example, moviegoers can examine this to determine whether a movie is worth watching. 

The other is a table from the-numbers.com that describes the market share for movie genres between 1995-2020. This dataset is more useful, for example, to people in the movie industry who are coming up with a new concept for a movie. It could also be useful for movie marketing teams to come up with a good strategy that could result in higher box office numbers and/or ratings

# Set Up

### Clean up environment and load packages
```{r}
# clean up the RStudio environment 
rm(list = ls())

# load all packages here: `mosaic`, `tidyverse`, and all others used
library(mosaic)
library(tidyverse)
library(tidyr)
library(rvest)
library(lubridate)
library(ggplot2)

#to import dataset I need
library(readr)

```

### Load the data you need
```{r}
Movie_Ratings <- read_csv("Movie Ratings.csv")
View(Movie_Ratings)
```

```{r}
#look at the first 6 rows
head(Movie_Ratings, 6)
```
Movie_ratings (from Kaggle) has the following case definition: "list the budget, audience, and critic ratings for a given movie".

```{r}
#rename columns for readability
Movie_Ratings <- Movie_Ratings %>% 
  rename(critic = 'Rotten Tomatoes Ratings %',
         audience = 'Audience Ratings %',
         year = 'Year of release')
```

The second data source comes from 'the-numbers.com'. However, I will be using a cleaned version of the data, which is loaded in a csv format.

```{r}
Genres <- read.csv("C:/Users/18vde/OneDrive/Desktop/PSU/STAT 184/Genres.csv")

Genres <- Genres %>% rename (rank = 'ï..Rank',
                             'BO_Rev (millions)'  = 'Total.Box.Office')

head(Genres, 6)
```
From this table, we can see that the case meaning is the following: "list the number of movies, box office revenue (in millions), number of tickets sold, and the market share (%) for each genre".


### Similarities between the data?

We can see that Movie_Ratings and Genres have a common variable: Genre. But how many values are common between the two data frames?
```{r}
Genres %>% distinct(Genre)
```

```{r}
Movie_Ratings %>% distinct(Genre)
```
As we can see, each genre from Movie_Ratings is present, which can lead us to ask the research question presented in the next section.

Let's look at these two data frames joined.
```{r}
MovieData <- Movie_Ratings %>% left_join(Genres, by = "Genre")

head(MovieData, 6)
```
The case meaning here is a bit different: not only does it look at the ratings (critic and audience) as well as budget for a given movie, but it also looks at how the film's genre has been doing for the past 25 years (examine the number of movies made and tickets sold, the overall box office revenue, and the overall ranking of the genre).

Now let's look at a summary for each
```{r}
summary(MovieData)
```
# Research Question

After examining the data, a good research question that we could explore would be the following:

**Does the market share of a genre have any influence on audience and critic ratings for a particular film?**

While the data frames show little evidence of maintenance (Movie_Ratings was uploaded to Kaggle 2 years ago; Genres was only published recently), they are still very useful to answer this question. 

It may not always be the case that market share numbers will result in a film being highly rated by viewers. For example, a film may make higher profits and its genre may have a large market share, but critics often disagree with viewers due to oversaturation of film ideas. There is also the case of people who are "ride or die" fans of a given genre. Although ratings can be very subjective, there are often large groups of people who push the profits of a certain genre to the higher end. 


# Exploring RQ

Since my research question looks at market share audience ratings, and critic ratings, I will use the following variables:

* audience
* critic
* Share

we'll need to understand which genres are most preferred and which make the most revenue in order to complete this question.

```{r}
#Explore distribution for audience ratings
MovieData %>% ggplot(aes(x = audience)) +
  geom_density(aes(color = Genre))
```

Overall, Thriller, Romance, and Drama movies are among the highest-rated films by audience members, but what about critics?

```{r}
#Explore distribution for critic ratings
MovieData %>% ggplot(aes(x = critic)) +
  geom_density(aes(color = Genre))
```

Critics agreed with audiences when it came to thrillers and Romance films, but at lower levels than audiences. They also considered adventure movies to be higher rated than drama films. In order to make sense of the research question, however, we'll need to examine the difference between critic and audience scores.

```{r}
# find difference between audience and critic
MovieData %>% mutate(difference = abs(audience - critic))
```

Since we are also looking at the market share per genre, we'll need to understand how the market share is distributed among the movie industry. We might get a clearer understanding of this by using the 'Genres' table, as that has a clearer breakdown of the market share

```{r}
Genres %>% ggplot(aes(x = Share, y = Genre)) +
  geom_col()
```