---
title: "Final Project EDA"
author: "Vaishali Devarakonda"
date: "Nov. 12, 2020"
output: html_notebook
---

#Intro

I will be examining the two data sources for this project: the first source is a dataset from Kaggle describing audience and critic ratings for a specific movie, created by user Mormukutchaudhary. The other is a table from the-numbers.com that describes the market share for movie genres between 1995-2020.

# Set Up

### Clean up environment and load packages
```{r}
# clean up the RStudio environment 
rm(list = ls())

# load all packages here: `mosaic`, `tidyverse`, and all others used
library(mosaic)
library(tidyverse)
library(tidyr)
library(rvest)
library(lubridate)
library(ggplot2)

#to import dataset I need
library(readr)

```

### Load the data you need
```{r}
Movie_Ratings <- read_csv("Movie Ratings.csv")
View(Movie_Ratings)
```

```{r}
#look at the first 6 rows
head(Movie_Ratings, 6)
```
Movie_ratings (from Kaggle) has the following case definition: "list the budget, audience, and critic ratings for a given movie".

```{r}
#rename columns for readability
Movie_Ratings <- Movie_Ratings %>% 
  rename(critic = 'Rotten Tomatoes Ratings %',
         audience = 'Audience Ratings %',
         year = 'Year of release')
```

For the second data source, we'll need to extract it from 'the-numbers.com'

```{r}
url <- "https://www.the-numbers.com/market/genres"

tableList <- url %>%
  read_html() %>%
  html_nodes(css = "table") %>%
  html_table(fill = TRUE)

Genres <- tableList[[1]]
head(Genres)
```
Genres (from the-numbers.com) has the following case definition: "list the number of movies, total box office revenue, number of tickets sold, and the total market share for each movie genre".

However, we'll need to get some cleaning done in order to make this data frame useful.

```{r}
Genres <- Genres %>% mutate(Tickets = as.integer(Tickets),
                            Movies = as.integer(Movies))
```
